Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-13 Final Draft/Final-Draft.tex
wording

\author{
  Catherine Darin\\
  Zagreb Mukerjee\thanks{
  Darin: Harvard University, \href{mailto:cdarin@hks.harvard.edu}{cdarin@hks.harvard.edu}.
  Mukerjee: Harvard University, \href{mailto:zagrebmukerjee@fas.harvard.edu}{zagrebmukerjee@fas.harvard.edu}.
  Code available at https://github.com/zagrebmukerjee/ReplicationPaper
}
}

@@

Like \cite{Baccini21}, our results suggest that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. However, our results differ from
\cite{Baccini21} in regards to the \emph{timing} of the impact of
manufacturing job losses on employment. In fact, using our preferred
measure of deindustrialization - \emph{net} change in manufacturing
employment - we find that counties with manufacturing job \emph{gains}
from 2012-2015 tended to swing towards Donald Trump in 2016. Given this
counter-intuitive result, we extend the analysis to the 2004-2015 period
in order to capture the long-term effects of deindustrialization on a
region. Doing so, we find that regions experiencing more manufacturing
job losses from 2004-2015 swung towards Donald Trump in 2016, and that
this occurred more when those losing jobs were white. We get similar
result using our covariate-matching approach. Relative to
\cite{Baccini21}, our results point to the substantive importance of the
timing of manufacturing job losses on political outcomes.

@@

\caption{\label{tab:descTable}Manufacturing Job Changes 2004-2015}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrr}
\toprule
  & Mean & Std Dev & 25th Pctile & Median & 75th Pctile\\
\midrule
Mfg Share of Emp & 0.20 & 0.15 & 0.09 & 0.17 & 0.29\\
Mfg Share of Emp (White) & 0.16 & 0.13 & 0.06 & 0.13 & 0.22\\
Mfg Share of Emp (Nonwhite) & 0.05 & 0.07 & 0.01 & 0.02 & 0.06\\
Change in Mfg Jobs/ Worker & 0.04 & 0.09 & -0.01 & 0.02 & 0.06\\
Change in Mfg Jobs/ Worker (W) & 0.03 & 0.07 & 0.00 & 0.02 & 0.06\\
\addlinespace
Change in Mfg Jobs/ Worker (NW) & 0.00 & 0.04 & -0.01 & 0.00 & 0.01\\
\bottomrule
\end{tabular}}
\end{table}
\FloatBarrier

@@

A potential weakness of our first approach is that manufacturing
employment in a county may be correlated with swinging towards Trump for
reasons other than manufacturing job losses. To account for this
potential ommitted variable bias, we use a second approach that attempts
to balance the covariates of the deindustrialization treatment. We use
the same covariates as are used in controls above, as well as
manufacturing as share of the population and total population. We apply
the method of Covariate Balancing Propensity Scores, as developed by
\cite{Imai14}; this method generates weights to reduce the correlation
between covariates and treatment, while addressing several of the issues
with propensity score weights. We find significant imbalance in the data
- in other words, it is relatively easy to predict the `treatment' of
job loss given the covariates (see Appendix for balance plots). Given
covariate-balancing weights, we then conduct a single-stage weighted
regression of change in Democratic vote share on manufacturing job loss,
using the same controls as above. Essentially, this approach can be
conceptualized as a form of matching that reduces the potential ommitted
variable bias in our first approach, but may introduce selection bias
(as some counties will be given low weights).

>>>>

\author{
  Catherine Darin\\
  Zagreb Mukerjee\thanks{
  Darin: Harvard
University, \href{mailto:cdarin@hks.harvard.edu}{cdarin@hks.harvard.edu}.
  Mukerjee: Harvard
University, \href{mailto:zagrebmukerjee@fas.harvard.edu}{zagrebmukerjee@fas.harvard.edu}.
  Code available at https://github.com/zagrebmukerjee/ReplicationPaper;
  https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/7U2UIQ
}
}

@@

Like \cite{Baccini21}, our results suggest that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. However, our results differ from
\cite{Baccini21} in regards to the \emph{timing} of the impact of
manufacturing job losses on employment. In fact, using our preferred
measure of deindustrialization - \emph{net} change in manufacturing
employment - we find that counties with manufacturing job \emph{gains}
from 2012-2015 tended to swing towards Donald Trump in 2016. Given this
counter-intuitive result, we extend the analysis to the 2004-2015 period
in order to capture the long-term effects of deindustrialization on a
region. Doing so, we find that regions experiencing more manufacturing
job losses from 2004-2015 swung towards Donald Trump in 2016, and that
this occurred more when those losing jobs were white. We get
directionally similar results using our covariate-matching approach -
differences in magnitude are an object of further investigation.
Relative to \cite{Baccini21}, our results point to the substantive
importance of the timing of manufacturing job losses on political
outcomes.

@@

\caption{\label{tab:descTable}Manufacturing Job Changes 2004-2015}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrr}
\toprule
  & Mean & Std Dev & 25th Pctile & Median & 75th Pctile\\
\midrule
Mfg Share of Emp & 0.13 & 0.10 & 0.05 & 0.11 & 0.19\\
Mfg Share of Emp (White) & 0.10 & 0.08 & 0.04 & 0.08 & 0.15\\
Mfg Share of Emp (Nonwhite) & 0.03 & 0.04 & 0.00 & 0.01 & 0.04\\
Mfg Job Loss/ Worker & 0.02 & 0.06 & 0.00 & 0.01 & 0.04\\
Mfg Job Loss/ Worker (W) & 0.02 & 0.04 & 0.00 & 0.01 & 0.04\\
\addlinespace
Mfg Job Loss/ Worker (NW) & 0.00 & 0.03 & 0.00 & 0.00 & 0.00\\
\bottomrule
\end{tabular}}
\end{table}
\FloatBarrier

@@

A potential weakness of our first approach is that manufacturing
employment may be correlated with swinging towards Trump for reasons
other than manufacturing job losses. To account for this potential
ommitted variable bias, we use a second approach that attempts to
balance the covariates of the deindustrialization treatment. We use the
same covariates as are used in controls above, as well as manufacturing
as share of the population and total population. We apply the method of
Covariate Balancing Propensity Scores, as developed by \cite{Imai14};
this method generates weights to reduce the correlation between
covariates and treatment, while addressing several of the issues with
propensity score weights. We find significant imbalance in the data - in
other words, it is relatively easy to predict the `treatment' of job
loss given the covariates (see Figures \ref{balPlot} and
\ref{balPlotWhite} for balance plots). Given covariate-balancing
weights, we then conduct a single-stage weighted regression of change in
Democratic vote share on manufacturing job loss, using the same controls
as above. Essentially, this approach can be conceptualized as a form of
matching that reduces the potential ommitted variable bias in our first
approach, but may introduce selection bias (as some counties will be
given low weights).
